---
title: "analysis_data"
output: html_document
---
The data set that will be use at this section will be dt2 (from processing_code file)
```{r}
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse)
library(tidymodels)  # for the parsnip package, along with the rest of tidymodels
library(readr)       # for importing data
library(lmtest) #assumptions validation
install.packages("broom.mixed")
library(broom.mixed) # for converting bayesian models to tidy tibbles
install.packages("dotwhisker")
library(dotwhisker)  # for visualizing regression results
library(MASS) # for box-cox transformation
library(recipes)
library(parsnip)
library(dials)
library(tune)
library(yardstick)
library(rsample)
library(workflows)
library(rpart)
library(rpart.plot)
library(glmnet)
library(ranger)
library(vip)
library(tree)
```
# Path to data
```{r}
data_location <- here::here("code","processing_code", "dt1.rds")
dt1 <- read_rds(data_location)
```
# Modeling
The model propose is a multiple linear regression model.
```{r}
#Outcome variable = COVID-19 and Pneumonia Deaths (pc_d)
#Independent variables = Sex, Group Age, State

#Build a Model

fit <- lm(pc_d~ age_group + State + Sex, data=dt1 )

lmtable<-summary(fit)
lmtable
table_file = here("products","graphs", "lmtable.rds")
saveRDS(lmtable, file = table_file)


layout(matrix(c(1,2,3,4),2,2)) 
plot(fit) 


#figure_file = here("products","graphs","modelplot1.png")
#ggsave(filename = figure_file, plot=modelplot1) 



```
It is observed then that in spite of having a relatively good coefficient of determination, there are problems related to the assumptions, and for this reason the box-cox transformation is used.

# Box-cox transformation

```{r}

bm=boxcox(as.numeric(dt1$pc_d) ~ dt1$age_group+ dt1$State + dt1$Sex )

lambda=bm$x[which(bm$y==max(bm$y))]

fit2 <- lm(pc_d^lambda~ age_group+ + State + Sex, data = dt1)

table_Boxcox<-summary(fit2)
table_Boxcox
table_file = here("products","graphs", "table_Boxcox.rds")
saveRDS(table_Boxcox, file = table_file)
```
After performing the box-cox transformation, the model is run again and an improvement in the coefficient of determination is observed and at the same time, graphically and analytically, it can be said that the assumptions of the model are being fulfilled (except for the autocorrelation assumption, but this could possibly be due to the nature of the data or the data collection). The assumptions were validated through tests such as shapiro wilks for normality, reset for linearity (although it is not very common to evaluate this assumption through an analytical test), Breusch-Pagan for the homogeneity of variance assumption and Durbin-Watson for the autocorrelation assumption.

# Assumption validation

```{r}
#lineality 
resettest(fit2)

# variance 
bptest(fit2)

# autocorrelation
dwtest(fit2,alternative = "two.sided")

# normality
shapiro.test(fit2$residuals)

```
# ANOVA

In this case, with a significance level of 0.05, non significant differences are observed between 
the age groups and state and sex.

```{r}
table_Anova<-anova(fit2)
table_Anova
table_file = here("products","graphs", "table_Anova.rds")
saveRDS(table_Anova, file = table_file)

```

#### MACHINE LEARNING ####
```{r}
machinedata <- dt1
colnames(machinedata)[10] <- "coviddeats"
colnames(machinedata)[9] <- "age"

set.seed(123) # 
sampledata <- sample(1:nrow(machinedata), size = nrow(machinedata)/2) # sample
trainset <- machinedata[sampledata,] # train data
testset <- machinedata[-sampledata,] # test data


set.seed(123) # seed
treeregresion <- tree::tree(
        formula = as.numeric(pc_d)~ as.factor(Sex)+as.factor(age_group), # Response variables and predictors variables
        data = trainset, # Data that will be used for the model creation 
        split = "deviance", # Deviation
        mincut = 20, # Minimum number of observations for division to occur
        minsize = 50 # Minimum number of observation for ramification
)

summary(treeregresion)

par(mar = c(1,1,1,1)) # plot margins
plot(x = treeregresion, type = "proportional") #  # Decision Tree (heterogeneity)
text(x = treeregresion, splits = TRUE, pretty = 0, cex = 0.8, col = "firebrick") #  Tree Label

regression_tree <-  tree::tree( 
                    formula = as.numeric(pc_d)~as.factor(Sex)+as.factor(age_group),
                    data    = trainset,
                    split   = "deviance",
                    mincut  = 1,
                    minsize = 2,
                    mindev  = 0
                  )

set.seed(500) 
cv_arbol <- cv.tree(regression_tree, K = 5) # cross crossing validation
cv_arbol


size <- rev(cv_arbol$size)[which.min(rev(cv_arbol$dev))] # Looking for the size
paste("Optimal size of the number of terminal nodes:", size)

final_tree <- prune.tree(
                  tree = regression_tree, #Tree size
                  best = size 
               )

predictions <- predict(regression_tree, newdata = testset)
test_rmse    <- sqrt(mean((predictions - testset$coviddeats)^2,na.rm=TRUE))
paste("Initial tree test error:", round(test_rmse,2))

predictions1 <- predict(final_tree, newdata = testset)
test_rmse1   <- sqrt(mean((predictions1 - testset$coviddeats)^2, na.rm=T))
paste("Final tree test error:", round(test_rmse1,2))
```

