---
title: "processed_data"
output: html_document
---

In this section you will find the data exploration and clean-up for furute analysis.

```{r}
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse)
```

# Path to data
```{r}
data_location <- here::here("Data","raw_data","Provisional_COVID-19_Deaths_by_Sex_and_Age.csv")
rawdata <- read.csv(data_location)
```
# What is in my data?
```{r}
#attach(rawdata)
tableRawData<- skimr::skim(rawdata)
tableRawData
table_file = here("products","graphs", "tableRawData.rds")
saveRDS(tableRawData, file = table_file)

```

This exploration shows an important amount of missing data in numerical variables, such as COVID.19.Deaths, Total.Deaths, Pneumonia.Deaths, Pneumonia.and.COVID.19.Deaths, Influenza.Deaths, and Pneumonia..Influenza..or.COVID.19.Deaths. Is important analysis is the amount of missing values are associated with other factors, such us age, state, sex, and year, before take a decsion about those values.
```{r}
# 1. Clean-up and rename variables

dt<-rawdata %>%
          dplyr::select(-c(Data.As.Of, Footnote)) %>%
    
          rename(age_group = Age.Group, startdate= Start.Date, endate = End.Date, 
          COVID19_d =   COVID.19.Deaths, total_d = Total.Deaths, pneum_d = Pneumonia.Deaths,
          flu_d = Influenza.Deaths, pfc_d = Pneumonia..Influenza..or.COVID.19.Deaths,
          pc_d = Pneumonia.and.COVID.19.Deaths)
dt
```
```{r}
# 2. Plots

plot1<-ggplot(data= dt, mapping = aes(x= COVID19_d, y=age_group)) +
  geom_boxplot() #Shows that young ages didn't die due COVID-19
plot1
figure_file = here("products","graphs","plot1.png")
ggsave(filename = figure_file, plot=plot1) 

plot2<-ggplot(data= dt, mapping = aes(x= pneum_d, y=age_group)) +
  geom_boxplot() #Shows that young ages didn't die due Pneumonia
plot2
figure_file = here("products","graphs","plot2.png")
ggsave(filename = figure_file, plot=plot2) 


plot3<-ggplot(data= dt, mapping = aes(x= flu_d, y=age_group)) +
  geom_boxplot() #Shows less deaths compared with P
plot3
figure_file = here("products","graphs","plot3.png")
ggsave(filename = figure_file, plot=plot3) 

plot4<-ggplot(data= dt, mapping = aes(x= pfc_d, y=age_group)) +
  geom_boxplot() #Shows that young ages area didn't die due COVID-19
plot4
figure_file = here("products","graphs","plot4.png")
ggsave(filename = figure_file, plot=plot4) 

plot5<-ggplot(data= dt, mapping = aes(x= COVID19_d, y=State)) +
  geom_boxplot()  
plot5
figure_file = here("products","graphs","plot5.png")
ggsave(filename = figure_file, plot=plot5) 

plot6<-ggplot(data= dt, mapping = aes(x= COVID19_d, y=Sex)) +
  geom_boxplot() # There were more male deaths than female deaths
plot6
figure_file = here("products","graphs","plot6.png")
ggsave(filename = figure_file, plot=plot6)

table_year <- table(dt$Group, dt$Year )
table_year
table_file = here("products","graphs", "table_year.rds")
saveRDS(table_year, file = table_file)

```
This data set can be analyzed using totals or using categories, for example All Sexes or Male and Female, or By Month, Year or total. To answer the project question will be more beneficial no use totals.
```{r}
# 3. Filters 

dt1 <- dt%>% filter(!age_group == "All Ages" & !age_group == "0-17 years" & 
                   !age_group == "1-4 years" & !age_group == "15-24 years" & 
                   !age_group == "5-14 years" & !age_group == "Under 1 year" & 
                   !age_group == "18-29 years")
                     

dt1 <- dt1%>% filter(!State == "United States")
dt1 <- dt1%>% filter(Group == "By Year")
dt1 <- dt1%>% filter(!Sex=="All Sexes") 
#dt1 <- dt1%>% filter(!pc_d == "NA")
dt1 <- dt1%>% filter(!pc_d== 0)
dt1 <- dt1%>% filter(pc_d >100)

#view(dt1)
skimr::skim(dt1)
```
#Save GA data 
```{r}
# The data will be saved as rds

# Data location 
save_data_location <- here::here("code", "processing_code","dt1.rds")

saveRDS(dt1, file = save_data_location )
```



  


